---
title: "Text Mining Saudi Aramco Prospectus"
author: "Rafael Nicolas Fermin Cota (nico)"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_depth: 4
    toc_float: 
      collapsed: TRUE
      smooth_scroll: FALSE
    number_sections: FALSE
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
rm(list = ls())
graphics.off()

# knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_chunk$set(
  collapse=TRUE,
  comment="#",
  message=FALSE,
  warning=FALSE,
  cache=TRUE, 
  # autodep = TRUE,
  fig.align = "center"
)

lightbox_img <- function(url, alt = "", caption = "", preview = TRUE) {
  if (preview) {
    glue::glue(
      '<a href="{url}" data-featherlight="image">
      <div class="figure">
      <img src="{url}" alt="{alt}">
      <p class="caption">{caption}</p>
      </div>
      </a>
      '
    )
  } else {
    if (alt == "") alt <- "static image of the plot"
    glue::glue('<a href="{url}" data-featherlight="image">{alt}</a>')
  }
}

```

# Extracting the prospectus text with pdftools

We used the [[pdftools]]{.pkg} package by [ROpenSci] to extract the text from the document, using the prospectus posted by Saudi Aramco, [available here](https://www.saudiaramco.com/-/media/images/investors/saudi-aramco-prospectus-en.pdf).
Extracting the text was as simple as downloading the PDF and running `pdftools::pdf_text()`.
We added page and line numbers to the extracted text and stored the result as a CSV.

```{r}
library(knitr)
library(ggpage)
library(pdftools)
library(magrittr)
library(tidyverse)
```

```{r}
path_root=file.path("..", "..")
# Download report from link above
aramco_prospectus_pdf <- pdf_text(
  file.path(path_root, "data", "saudi-aramco-prospectus-en.pdf")
)

aramco_prospectus_text <- tibble( # https://www.tidytextmining.com
  page = 1:length(aramco_prospectus_pdf),
  text = aramco_prospectus_pdf
) %>% 
  separate_rows(text, sep = "\n") %>% 
  group_by(page) %>% 
  mutate(line = row_number()) %>% 
  ungroup() %>% 
  select(page, line, text)

# write_csv(aramco_prospectus_text, file.path("data", "saudi-aramco-prospectus-en.csv"))
# head(aramco_prospectus_text) %>% kable
```

## Visualizing the prospectus pages with ggpage

Emil Hvitfeldt released [[ggpage]]{.pkg}, a package that lets you create a page-layout visualization using [[ggplot2]]{.pkg}.
While the package uses the text content of the document only --- so the visualized text layout doesn't completely match the layout of the original document --- it does allow you to highlight text elements.

The first step is to load the text version of the Saudi Aramco Prospectus. You can see from the first few lines of the data that the OCR really struggled with the header that appears at the top of each page and has been crossed out with a single line.
(The redacted text is less confusing to the OCR because it's rendered in solid black and generally results in blank space.)

The core of the next step is to pass the `aramco_prospectus_pages` to `ggpage::ggpage_build()`. Before doing that, though, we pad each page to make sure they have the same number of lines.

The `ggpage_build()` function tokenizes the text into individual `word`s, so we then use `str_detect()` to find mentions of the key players.

```{r ggpage_build}
aramco_prospectus_pages <- aramco_prospectus_text %>% 
  # pad pages with fewer lines than expected
  complete(
    page, 
    line = 1:max(aramco_prospectus_text$line),
    fill = list(text = "")
  ) %>% 
  # Pre-process for {ggpage}
  ggpage_build(
    ncol = 30, 
    bycol = FALSE, 
    page.col = "page", 
    wtl = FALSE, 
    x_space_pages = 10,
    y_space_pages = 100
  ) %>% 
  mutate(
    color = case_when(
      str_detect(word, "reserves|life") ~ "Reserves",
      str_detect(word, "gearing")     ~ "Gearing",
      str_detect(word, "roace")      ~ "Capital",
      TRUE ~ "risk"
    ),
    color = factor(color, c(
      "Reserves", "Gearing", "Capital", "risk"
    ))
  )
# aramco_prospectus_pages
```

The following bit of code sets up the color palette, which is derived from the `Libre Office Calc` theme provided by [[ggthemes]]{.pkg}.

```{r plot-colors}
# manually assigned colors from ggthemes::pal_calc()
colors <- rep("", length(levels(aramco_prospectus_pages$color)))
names(colors) <- levels(aramco_prospectus_pages$color)
colors["Reserves"]    <- "#FF4023"
colors["Gearing"]   <- "#004983"
colors["Capital"]    <- "#FF922E"
colors["risk"]   <- "#d0d0d0"
```

Finally, `ggpage_plot()` from [ggpage]{.pkg} creates the [ggplot2]{.pkg} page layout, and adding the fill aesthetic using the manual color scale defined above adds color highlights for mentions of Reserves, Gearing, and others.

```{r aramco-ggpage, fig.show = "hide", fig.width=18, fig.height=11, dependson="ggpage_build"}

# pdf(file="aramco_prospectus_pages.pdf")

ggpage_plot(aramco_prospectus_pages) +
  aes(fill = color) +
  scale_fill_manual(
    values = colors, 
    breaks = setdiff(names(colors), "risk")
  ) +
  labs(fill = NULL, caption = "@nico") +
  guides(fill = guide_legend(nrow = 1)) +
  theme(legend.position = "bottom")

# dev.off()
```

<a name="ggpage-plot"></a>
  
  `r lightbox_img(knitr::fig_chunk("aramco-ggpage", "png"))`

_Click the image to expand._

```{r, echo=FALSE}
# rm(aramco_prospectus_pdf, aramco_prospectus_text, aramco_prospectus_pages)
# rm(lightbox_img, colors)
```
