---
title: "Time Series and DCF Analysis in R"
author: "Karl Polen"
date: "9/20/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(warning = FALSE,message=FALSE)
library(shiny)
library(shinydashboard)
library(tidyverse)
library(readxl)
library(purrr)
library(lubridate)
library(xts)
library(timetk)
library(asrsMethods)
library(knitr)
library(DT)
library(leaflet)
library(plotly)
source("apt_functions.R")
```

## What we will cover in this presentation

- Some background discussion on techniques and goals for DCF analysis of real estate projects
- Techniques for doing cash flow analysis using the XTS time series package illustrated by a simple example of an investment in a rental house
- Presentation of a Shiny app that shows the analysis of the Modera Decatur project 


## What's inside the box?

Materials for this lecture are found at https://github.com/karlp-asrs/nus_apartment inside the apt_app folder

- This slide presentation and the RMD that generated it
- apt.function.R is a set of functions to analyze apartment development projects
- modera_decatur.xlsx is the configuration file for the Modera Decatur project (you can modify this file to change the parameters or create a completely new project)
- app.R is a shiny app that shows the results of the analysis and allows you to perform sensitivity to analysis to key risk factors

## DCF analysis for real estate development projects

- The purpose of real estate underwriting is to grasp the likelihood that a project will be successful and to understand risks that threaten that
- The DCF output summarizes the results of that analysis in a way that allows you to calculate financial metrics
- Let me repeat, the DCF is a necessary technique but it isn't the analysis.  The apparent 'scientificness' of DCF analysis can be a deceptive distraction.
- The real analysis happens before you start the DCF in understanding the market and competition for your project and the feasibility of the construction timeline and budgets.

## The risks in a real estate development project -- predevelopment phase

- The predevelopment process consists of obtaining necessary governmental approvals
- These approvals can result in a change in the scope of a project with potential impacts to both density and cost
- An experienced manager understands this process and should be expected to mitigate negative impacts of changes
- Delays are common as a result of these negotiations

## Construction phase risks

- During predevelopment, management will have completed plans for the project and obtained bids for the work
- Experienced managers should be expected to deliver projects on time and on budget
- Failures in this regard are typically the result of poor management and not indicative of the inherent risks and difficulty of construction
- Nevertheless, even with competent management external events such as natural disasters can impact supply chains causing delays
- A DCF analysis should consider the impacts of possible delays and cost overruns

## Leaseup phase risk

- The development process takes several years and there is inherent risk that markets will evolve by the time your project is ready for occupancy
- Your analysis should consider that rents and pace of leaseup may be different from the initial underwriting
- Model a balance sheet to show the viability of your financial structure under stress

## Capital markets risk

- Evolving capital markets can impact your project
- The key phase of risk is in transition from construction financing to permanent financing
- If a recession hits while your project is delivered to the market, this is a challenge and one of the bigger sources of bankruptcies in real estate
- Analysis doesn't help solve this.  Sponsors and investors need to work collaboratively to deal with this dynamic if it happens in order to avoid destruction of capital.
- Use leverage prudently and maintain appropriate reserves

## DCF analysis in R -- the basics

- Use `xts` time series objects to record cash flow amounts and dates.  For example, you buy an asset for \$100 and sell it for \$150 18 months later.

```{r}
buydate=as.Date("2016-01-01")
(buy=xts(-100,buydate))
(sell=xts(150,buydate+months(18)))
```

## Time series arithmetic

- you can do calculations on xts time series, but xts requires a match on the time attribute.  For example, `buy+sell` produces `zilde` because there is no match in the date.

```{r}
buy+sell
```

On the other hand, this works, sort of. . .

```{r}
sum(buy,sell,na.rm=TRUE)
```

## Assembling time series in a matrix

- The "go-to" technique is to assemble your time series as a matrix with `cbind` and then use `rowSums`.  I use this technique a lot because it allows you to assemble and visualize data in logical subsets.  It also allows you to deal with data of indeterminate length as illustrated with the following code. 

```{r}
cflist=list(Buy=buy,Sell=sell)
cfmatrix=do.call(cbind,cflist) 
cfmatrix=cbind(cfmatrix, xts(rowSums(cfmatrix,na.rm=TRUE),index(cfmatrix)))
names(cfmatrix)=c(names(cflist),"Total_CF")
cfmatrix
```

## Analyzing a rental house investment

- I will show code to analyze a rental house using the XTS package in R
- Here are key assumptions:
  * Buy the house for 300,000 with a 250,000 mortgage at 3% with 30 yr amortization
  * Renovation of 20,000 before you start leasing
  * Rent is 20,000 per year and increases 4% per year
  * Maintenance is weekly pool and landscape service at 50 per week
  * Tenant pays utilities
  * Insurance is 2000 paid annually at closing and each anniversary
  * Property taxes are 3,000 paid semi-annually in May and November
  * Market value of rental houses is 15 times annual rent
  
## Buy the house

```{r}
start_date=as.Date("2020-08-01")
house_purchase_price=300000
buy_house=xts(-house_purchase_price,start_date)
mortg_proceeds=xts(250000,start_date)
#loanamort is a function for loan calcs from asrsMethods
mortg_vars=loanamort(r=.03,bal0=250000,n=30,freq=12,apr=TRUE)  
mort_dates=start_date+months(0:360)
mortg_bal=xts(c(250000,mortg_vars$bal),mort_dates)
mortg_int=xts(c(0,-mortg_vars$int),mort_dates)
mortg_pmt=xts(c(0,rep(-mortg_vars$pmt,360)),mort_dates)
```
  
## Renovate it

```{r}
renovation_amt=20000
renovation=xts(rep((-renovation_amt/6),6),start_date+months(1:6))
```

## Lease it

```{r}
market_rent=(20000/12)*exp(cumsum(c(0,rep(log(1+(.04/12)),360)))) 
market_rent=xts(market_rent,start_date+months(0:360))
annual_rent=market_rent[6+(12*(0:11))]
monthly_rent=rep(annual_rent,rep(12,length(annual_rent)))
monthly_rent=xts(monthly_rent,start_date+months(5)+
                   months(1:length(monthly_rent))+
                   days(5))  #note rent received on 5th day of month
```

## Expenses

```{r}
utilities_renovation=xts(rep(-300,6),start_date+months(1:6))
weekly_maint=xts(rep(-50,12*52),start_date+days(7*1:(12*52)))
insurance=xts(rep(-2000,12),start_date+years(0:11))
first_tax_date=ymd(paste0(year(start_date),"-05-15"))+months(c(0,6))
first_tax_date=first_tax_date[first_tax_date>start_date]
property_tax=xts(rep(-1500,24),first_tax_date[1]+months(6*(0:23)))
```              


## Tax Calcs
``` {r}
# house gets depreciation tax shield with useful life of 30 years
bldg_depreciation=xts(rep(-house_purchase_price/360,360),
                      start_date+months(0:359))
# renovation get depreciation tax shield with useful life of 5 years
renov_depreciation=xts(rep(-renovation_amt/60,60),
                       start_date+months(6)+months(1:60))
```

## Calculate value of house 
```{r}
house_value=market_rent*12*15
```

## Create a convenience function for displaying annual statements

```{r}
show_annual=function(xtsmat,nyear=6,isbs=FALSE) {
  xtsmat=apply(xtsmat,2,nona) # get rid of NAs
  df=as.data.frame(xtsmat)
  xtsdates=as.Date(rownames(df))
  rownames(df)=NULL
  years=year(xtsdates)
  year1=years[1]
  lastyear=tail(years,1)
  yearlabel=year1:lastyear
  if (isbs) {
    dfy=aggregate(df,by=list(years),FUN=lastinvec)} else {
    dfy=aggregate(df,by=list(years),FUN=sum)  }
  dfy$Group.1=NULL
  dfy=data.frame(Year=yearlabel,dfy)
  return(round(dfy[1:nyear,]))
}
```

## Show operating cash flow

```{r}
opcfmat=cbind(monthly_rent,utilities_renovation,
              weekly_maint,insurance,property_tax)
total_opcf=xts(rowSums(opcfmat,na.rm=TRUE),index(opcfmat))
opcfmat=cbind(opcfmat,total_opcf)
colnames(opcfmat)=c("Rent","Utilities","Landscape/pool","Insurance",
                    "Property_tax","Total_OPCF")
show_annual(opcfmat)
```


## Show total cash flow

```{r}
cfmat=cbind(total_opcf,buy_house,mortg_proceeds,mortg_pmt,renovation)
total_cfmat=rowSums(cfmat,na.rm=TRUE)
cfmat=cbind(cfmat,total_cfmat)
colnames(cfmat)=c("Op_CF","House_pur","Mort_proc",
                  "Mort_pmt","Renov","Total")
show_annual(cfmat)
```

## Show taxable income

```{r}
taxmat=cbind(total_opcf,mortg_int,bldg_depreciation,renov_depreciation)
total_tax=rowSums(taxmat,na.rm=TRUE)
taxmat=cbind(taxmat,total_tax)
colnames(taxmat)=cbind("Operating_CF","Mort_Interest","Bldg_depr",
                       "Reno_depr","Txbl_Inc")
show_annual(taxmat)
```

## Show balance sheet and growth of net worth

```{r}
owner_eq=cbind(house_value,-mortg_bal)
owner_eq=rowSums(owner_eq,na.rm=TRUE)
bsmat=cbind(house_value,mortg_bal,owner_eq)
colnames(bsmat)=c("House_Value","Loan_Balance","Owner_Equity")
show_annual(bsmat,isbs=TRUE)
```

## calculate IRR after 5 years

```{r}
end_date=start_date+years(5)
totalcf=cfmat[,"Total"]
irrcf=totalcf[index(totalcf)<=end_date]
irrval=bsmat[,"Owner_Equity"][end_date]
irrcfmat=cbind(irrcf,irrval)
irrcf_w_val=xts(rowSums(irrcfmat,na.rm=TRUE),index(irrcfmat))
irr.z(irrcf_w_val)
```

## Observations and recommendations

- Record the cash flows using their known patterns as daily events
- Maintain that detail for calculations, but summarize into calendar periods for presentation
- Use detail where you have it, but don't invent it when you are guessing
- Use broad categories for allowances and estimates
- The best model is the briefest model that accomplishes its purpose

## Present results in standard financial statement format

- This will facilitate communication by conforming to reporting norms
- It will facilitate comparison of underwriting forecasts to actual results
- The cash flow is necessary for investment metrics
- You need a balance sheet to assess viability of financial structure
- Taxable investors need an income statement to understand tax impacts
- Institutional investors need an income statement to decompose performance between operating income and appreciation

## The Modera Decatur project

A shiny app has been provided to illustrate this analysis

## Additional resources

This code uses the `asrsMethods` package which is available at github. The following code will install it. 

`devtools::install_github("karlp-asrs/asrsMethods/asrsMethods")`

Much more information on analyzing real estate and private equity is found at my github site https://karlp-asrs.github.io/

