---
title: "Extracting Data from the Saudi Arabia Capital Allocation Database"
author: "Nico"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: true
      smooth_scroll: true
    number_sections: true
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  # echo = TRUE, 
  collapse=TRUE,
  comment="#",
  message=FALSE,
  warning=FALSE,
  dev.args = list(pointsize = 11)
)

library(distill)
library(rmarkdown)

format.dt.f = function(
  df, 
  page_length = 10,
  perc_vars=NA,
  ron_vars=NA,
  ron_digits=2
){
  if( is.null(df) | purrr::is_empty(df) ){return()}
  
  double.two.int.f = function( df ){
    get_no_digits = function(x){
      if( ! is.numeric(x) ){return(NULL)}
      x = x %% 1
      x = as.character(x)
      no_digits = nchar(x) - 2
      no_digits = ifelse( no_digits == -1, 0, no_digits )
      return(no_digits)
    } 
    suppressWarnings({
      new_df = df %>%
        as_tibble() %>%
        mutate_if( function(x) max( get_no_digits(x), na.rm = T ) == 0, as.integer )
    })
    return(new_df)
  }
  df = double.two.int.f( df )
  max_length = nrow(df)
  page_length_menu = c(10,25,50,100, max_length, page_length) %>% unique()
  page_length_menu = page_length_menu[ !page_length_menu > max_length]
  
  dt = DT::datatable(
    df, 
    extensions = c('Buttons', 'ColReorder', 'KeyTable', 'FixedColumns'), 
    rownames = FALSE, 
    options = list(
      dom = 'Bflrtip', 
      buttons = I( c('colvis','copy', 'excel') ), 
      colReorder = TRUE, 
      keys = TRUE, 
      pageLength = page_length, 
      lengthMenu = page_length_menu,
      scrollX = TRUE,
      scrollCollapse = TRUE
    )
  )
  
  if (!is.na(ron_vars)[1]) dt=dt %>% DT::formatRound( ron_vars, ron_digits )
  if (!is.na(perc_vars)[1]) dt=dt %>% DT::formatPercentage( perc_vars, 2 )
  
  return(dt)
}
```

# Context

In this lecture, we look at how we can use R to automate the process of <a href="https://rpubs.com/rafael_nicolas/saudi_arabia_capital_allocation" target="_blank">extracting relevant fundamental data from the Saudi Arabia capital allocation database</a>.

# Initialization

We begin by: 

## Clearing all pre-existing objects from the workspace

```{r, eval=FALSE}
gc(); rm(list=ls()); graphics.off()
```

## Initializing the file paths

```{r}
path_root <- "." # Sets the current working directory as the relative file path
path_output <- file.path(path_root, "output") 
path_data <- file.path(path_root, "data") # ./data
```

Download the xlsx files here: <a href="https://github.com/rnfermincota/academic/tree/main/research/traditional_assets/database/data/saudi_arabia" target="_blank">https://github.com/rnfermincota/academic/tree/main/research/traditional_assets/database/data/saudi_arabia</a>

## Loading the necessary libraries

```{r, warning=FALSE, message=FALSE}
library(tidyxl)   # Imports non-tabular data from Excel files.
library(tibble)   # Stricter checking and better formatting than the traditional dataframe.
library(dplyr)    # A manipulation tool for working with data frame like objects.
library(purrr)    # A Complete functional programming toolkit for R.
library(furrr); plan(multisession) # Provides multicore parallel implementations of the purrr map() function. 
library(unpivotr) # Tools for converting data from complex or irregular layouts to a columnar structure. https://nacnudus.github.io/unpivotr/
```


# The `extract.sheets.f` Function

## What?

A function for extracting all information contained in all sheets in a specified Excel file. 
The `excl` parameter can contain a list of columns that should be excluded from being converted to _numeric_ type before the final dataframe is returned.


```{r}
extract.sheets.f <- function(dsheet, excl=NULL){
  
  out <- dsheet %>% 
    select(row, col, data_type, character, numeric, logical) %>%      # 1
    rectify %>%                                                       # 2
    select(-c("row/col"))                                             # 3
  
  nm <- out %>% slice(1) %>% unlist(use.names=FALSE) # headings       # 4
  out <- set_names(out, nm) %>% slice(2:n())                          # 5
  
  if (!is.null(excl)){                                                # 6
    out <- out %>% mutate(across(-excl, as.numeric))
    }
  
  return(out)                                                         # 7
}
```

## How?

  1. Filter: use the `select()` function to extract the relevant columns from the Excel sheet.
  2. Format: use the `rectify()` function to display the cells as though in a spreadsheet. _See example:_ <a href="https://github.com/nacnudus/unpivotr/blob/main/R/rectify.R" target="_blank">https://github.com/nacnudus/unpivotr/blob/main/R/rectify.R</a>

  3. Remove: use the `select()` function to remove the _row/col_ column.
  4. Headings: Extract and save the first row (currently storing the column headers) as `nm` using the `slice()`function. The `unlist(use.names=FALSE)` function is used to extract the column names as an atomic vector for easier manipulation.
  5. Set names: Set the column names of the `out` dataframe to our vector `nm` and at the same time, remove the first row that contains the initial column names.
  6. IF: A conditional statement for excluding previously specified columns from being converted to `numeric`. The `across()` function allows us to transform several column at once, instead of doing it manually, one by one.
  7. Return: Returns the final dataframe.


# Extract File Names

We use the `list.files()` function to extract the file names within our _"./data/xlsx"_ folder into a character vector.

```{r}
list_paths <- list.files(
  path = file.path(path_data),
  pattern = ".xlsx",
  full.names = TRUE
  )
```

We provide the following arguments:

  1. `path`: Directs the function to look within the previously specified _"./data/xlsx"_ directory.
  2. `pattern`: Tells the function to pick out all Excel files within the specified folder xlsx.
  3. `full.names`: Prepends the specified directory path to the file names.
  

```{r}
as.data.frame(list_paths) %>% 
  format.dt.f(.)
```


# Sheet Extraction Prep

## Create list of sheet names


```{r}
dat = enframe(list_paths, name = NULL, value = "path") %>%
      mutate(sector=gsub(".xlsx", "", basename(path)))
```

First, we use the `enframe()` function to convert the `list_paths` character vector into a one column dataframe since `name` parameter is set to `NULL` (otherwise, an additional column of names is also created). The `value` parameter sets the name of the single column returned. 

Next, the `mutate()` function is used to create a new column named `sector`. The `sector` column is created using the `gsub()` function, which uses a regular expression pattern to locate and _substitute_ the filename extension _.xlsx_ for an empty string. The `basename()` function is used to remove the prepended file path from each character string up to and including the last path separator.

The table below shows a _before_ and _after_ effect of this function in its two columns respectively:

```{r}
dat %>% 
  format.dt.f(.)
```


## Extract sheets and store in column

```{r}
dat = dat %>% 
      mutate(dataset = future_map(path, xlsx_cells) %>%
      set_names(sector))
```

Finally, we mutate the dataframe further by adding a new column called `dataset` which contains all data from all sheets contained  within the excel file specified in the corresponding `path` column. The `xlsx_cells()` function is used to import the data cell by cell in conjunction with the `future_map()` function which applies a particular function to each element of a vector in parallel. The extracted sheets are stored as `tibbles` in the `dataset` column.


# Sheet Extraction

## What?

In this section, we combine the `extracts.sheets.f()` function with the `dat` dataframe to extract all relevant data from the _earnings_debt_ sheets into a single dataframe.

## How?

First, we define a vector of column names as the variable `excl` for use in the `extracts.sheets.f()` function.

```{r}
excl <- c("country", "company_name", "industry_group")
```

Next,

```{r}
saudi_arabia_industries <- 
    future_map(dat$dataset, function(inp){  # 1
               extract.sheets.f(dplyr::filter(inp, sheet=="earnings_debt"), all_of(excl)) %>%
               slice(1)                     # 2
    }) %>%
    bind_rows(.) %>%                        # 3
    arrange(industry_group) %>%             # 4
    arrange(desc(roic_cost_capital))        # 5
```

1. Use the `future_map()` function to iterate over the `dat` dataframe, extracting all sheets named _earnings_debt_ stored in the `dataset` column (remember that this column stores `tibbles`). We use the `all_of()` function to make sure that every value in the `excl` vector is present.
2. Use the `slice()` function to grab only the first row at this time.
3. Since step (1) returns a list of `r nrow(saudi_arabia_industries)` elements, we use the `bind_rows(.)` function to bind the individual rows into a single dataframe which shows _earnings_debt_ data for all industries included in the _"./data/xlsx"_ folder.
4. Next, arrange the data alphabetically by `industry group` in ascending order. 
5. Finally, arrange the industries based on their `roic_cost_capital` in descending order.

A sample of the `saudi_arabia_industries` dataframe can be viewed

```{r}
saudi_arabia_industries %>% 
  format.dt.f(.)
```

In this next chunk of code, we perform the same steps as above except instead of only grabbing the first row from the `earnings_debt` sheet, we grabbed the remaining rows.

```{r}
saudi_arabia_earnings_debt <- 
    future_map(dat$dataset, function(inp){
               extract.sheets.f(dplyr::filter(inp, sheet=="earnings_debt"), all_of(excl)) %>%
               slice(2:n())
  }) %>%
  bind_rows(.) %>%
  arrange(industry_group) %>%
  arrange(industry_group, desc(roic_cost_capital))
```


```{r}
format.dt.f(saudi_arabia_earnings_debt)
```


# Dividend Yield Filter

1. Starts by selecting only rows from the `saudi_arabia_earnings_debt` dataframe that have a `dividend_yield` greater than $0.05$.
2. Groups the resulting filtered rows by `industry_group` and uses the `slice()` function to return only the first occurring result for each industry.

```{r}
saudi_arabia_earnings_debt %>%
  dplyr::filter(dividend_yield>0.05) %>%
  group_by(industry_group) %>%
  slice(1) %>% 
  format.dt.f(.)
```


# Capital Structure

The code here is very similar to the __Sheet Extraction__ section, but instead we target the `cost_capital` sheets for extraction

```{r}
excl <- c("company_name", "exchange_ticker", "industry_group", "country",
       "reported_debt_rating", "current_debt_rating", "optimal_debt_rating", 
       "flag_bankruptcy", "flag_refinanced")

saudi_arabia_cost_capital <- 
  future_map(dat$dataset, function(inp){
    if((inp$sheet %>% n_distinct) > 1){                                           # 1
        extract.sheets.f(dplyr::filter(inp, sheet=="cost_capital"), all_of(excl)) %>%
        mutate(
          flag_bankruptcy=as.logical(flag_bankruptcy),
          flag_refinanced=as.logical(flag_refinanced)
        )
    }
  }) %>%
  bind_rows(.) %>%
  mutate(spread_optimal = current_debt_capital-optimal_debt_capital) %>%           # 2
  arrange(industry_group, spread_optimal)
```

There are two notable exceptions:

1. The sheet is only loaded `if` the corresponding tibble in the `dataset` column has more than one distinct sheet.
2. An additional column `spread_optimal` is created by subtracting `optimal_debt_capital` from `current_debt_capital`.

```{r}
format.dt.f(saudi_arabia_cost_capital)
```



# Putting It All Together

In this section, we combine the data from our `saudi_arabia_earnings_debt` and `saudi_arabia_cost_capital` dataframes into one final dataframe before filtering out any rows that either exceed or fall short of certain metrics.  

```{r}
saudi_arabia_screener <- saudi_arabia_earnings_debt %>%
  select(                                                                              # 1
      industry_group, company_name, dividend_yield, roe, 
         cost_equity, roe_excess_return=roe_cost_equity, roic, 
         cost_capital, roic_excess_return=roic_cost_capital) %>%
  left_join(                                                                           # 2
    saudi_arabia_cost_capital %>% 
  select(company_name, current_debt_capital,                                            
         optimal_debt_capital, spread_optimal)) %>%
  select(company_name, dividend_yield, roic_excess_return, roic,                       # 3
         cost_capital, roe_excess_return, roe, cost_equity, spread_optimal, 
         current_debt_capital, optimal_debt_capital) %>%
  dplyr::filter(dividend_yield > 0.01, roic_excess_return > 0.025, spread_optimal < 0) # 4
```

1. Use the `select()` function to grab, and simultaneously rename, the columns of interest in the `saudi_arabia_earnings_debt` dataframe
2. Perform a `left_join()` with the `saudi_arabia_cost_capital` dataframe after selecting four columns of interest. Since no column to join the two dataframes as specified, the common column `company_name`  is used.
3. Once again, use the `select()` function to keep only the columns of interest
4. Use the `filter()` function to keep only the rows that pass ALL of the following conditions:
    
    * `dividend_yield` > 0.01
    * `roic_excess_return` > 0.025
    * `spread_optimal` < 0

```{r}
format.dt.f(saudi_arabia_screener)
```

# Using the `extract.sheets.f` to extract all sheets from an Excel file

This function combines and utilizes many of the tools we have already seen. We take things further by however, by extracting ALL sheets from a given Excel file and grouping them together in a database like format.

```{r}
if (FALSE){
  saudi_arabia_db <- 
    future_map(dat$dataset,function(inp){
      excl = c("country", "company_name", "industry_group")
      out = list(earnings_debt=extract.sheets.f(
                 dplyr::filter(inp, sheet=="earnings_debt"), all_of(excl)))       # 1
      
      if ( (inp$sheet %>% n_distinct) > 1){                                       
        excl = c("company_name", "exchange_ticker", "industry_group", "country",
               "reported_debt_rating", "current_debt_rating", "optimal_debt_rating", 
               "flag_bankruptcy", "flag_refinanced")
      
        out = list_modify(                                                        
          out,
          cost_capital = extract.sheets.f(
            dplyr::filter(inp, sheet == "cost_capital"), all_of(excl))  %>%      # 2
            mutate(
              flag_bankruptcy=as.logical(flag_bankruptcy),
              flag_refinanced=as.logical(flag_refinanced)
            )
        )

        nm = inp %>%                                                              # 3
          select(sheet) %>%
          distinct(.) %>%
          dplyr::filter(!sheet %in% c("earnings_debt", "cost_capital")) %>%
          pull

        out = list_modify(                                                        # 4
          out,
          optimal_mix = map(
            nm,
            ~extract.sheets.f(dplyr::filter(inp, sheet == .x), NULL) %>%
              mutate(across(c(where(is.character), -debt_rating), as.numeric))
          ) %>% set_names(nm)
        )
      }
      
      out                                                                         # 5
    }
  )
  listviewer::jsonedit(saudi_arabia_db)
}
```

1. Just as we have done previously, extract the `earnings_debt` sheet from the `dataset` column. This time however, store the extracted sheet in a list named `out`.  
2. Next, extract the `cost_capital` sheet, and append it to our master list of sheets `out`.
3. Create a vector of distinct sheet names that are not `cost_capital` or `earnings_debt`. This step uses the `pull()` function to extract a list of the remaining sheets in the provided Excel file.
4. Then, for every sheet name that was collected in the previous step, use the `extract.sheets()` function to extract the sheet corresponding to that name, storing it in the `optimal_mix` variable and appending all extracted sheets to our master list `out`.
5. Finally, Return the completed `saudi_arabia_db`. 


# Housekeeping

We finish by saving our four dataframes as an RDA file. RDA files allow us to export and store our work in a format that is shareable and restorable in any future R session.

```{r}
save(saudi_arabia_industries, saudi_arabia_earnings_debt, saudi_arabia_cost_capital, 
     saudi_arabia_screener, file=file.path(path_output, "saudi_arabia_fundamental_data.Rda"))
```

Finally, we remove all created objects and user-defined functions from the global environment.

```{r}
rm(dat)
rm(saudi_arabia_industries, saudi_arabia_earnings_debt, saudi_arabia_cost_capital, saudi_arabia_screener)
rm(path_root, path_data, path_output, list_paths)
rm(excl, extract.sheets.f)
```


# References

__Saudi Arabia: Capital Allocation via Cost of Capital__ <a href="https://rpubs.com/rafael_nicolas/saudi_arabia_capital_allocation" target="_blank">https://rpubs.com/rafael_nicolas/saudi_arabia_capital_allocation</a>